---
title: "UWWI Services cleanup"
output:
  html_notebook:
    code_folding: show
    toc: true
    toc_float: true
---

R code to take the services dataset and convert columns with lists of values, e.g. ['Sunday', 'Monday'] to long or wide format as desired.

The functions `getUniqueVals`, `replaceEmptyWithNA`, `stripEnclosingChars`, `spreadRowsToCols`, `splitListToCols`, should be usable with the other datasets.

```{r setup}
library(tidyverse)
library(foreach)

source("DataCleaningFunctions.R")

# Hide the dplyr .groups message
options(dplyr.summarise.inform = FALSE)
```

```{r}
services_original <- read_csv(file = "../uwwi_dataset_services.csv")
services <- services_original
```

```{r}
head(services)
```

```{r}
summary(services)
```

```{r}
services %>% head(200) %>% View()
```

# Exploration

Does the ServiceSystem_Active flag make sense?

```{r, rows.print = 50}
services %>%
  group_by(ServiceSystem_Active, ServiceOption_PermanentlyInactiveService) %>%
  summarise(n = n())
```

Some data cleaning issues.

# Data Cleaning

```{r}
# Function to strip start/end characters from a string
stripEnclosingChars <- function(s) {
  str_sub(s, 2, -2)
}

# Function to parse unique values from the list
getUniqueVals <- function(data, field) {
  # Unique list
  data %>%
    select({{field}}) %>%
    unique() %>%
    # Strip [ ]
    mutate(
      {{field}} := stripEnclosingChars({{field}})
    ) %>%
    # Split , into new rows
    separate_rows({{field}}, sep = ", ") %>%
    # Dedupe again
    unique() %>%
    # Strip ' '
    mutate(
      {{field}} := stripEnclosingChars({{field}})
    ) %>%
    return()
}
```

```{r}
services_original %>% getUniqueVals(ServiceOption_PaymentOptions) %>% print()
```

```{r}
# Function to replace empty string with NA
replaceEmptyWithNA <- function(s) {
  case_when(
    s == "" ~ "NA"
    , TRUE ~ s
  ) %>%
  return()
}

# Function to convert multi-row format to multi-column
spreadRowsToCols <- function(data, field, .unit = 1) {
    data %>%
    pivot_wider(
      names_from = {{field}}
      , values_from = .unit
      , values_fn = max
      , names_prefix = paste0(ensym(field), '_')
    ) %>%
    return()
}

# Function to split field into multiple rows
splitListToCols <- function(data, field, delimiter = ', ', cols = TRUE, .unit = 1) {
  df <- data %>%
    # Strip [ ]
    mutate(
      {{field}} := stripEnclosingChars({{field}})
    ) %>%
    # Split , into new rows
    separate_rows({{field}}, sep = ', ') %>%
    # Strip ''
    mutate(
      {{field}} := stripEnclosingChars({{field}}) %>% replaceEmptyWithNA()
      , .unit = .unit
    )
  
  if (cols) {
    df %>%
      spreadRowsToCols({{field}}, .unit) %>%
      return()
  } else {
    return(df)
  }
}
```

```{r}
# Unit tests
test <- services_original %>%
  head(20) %>%
  select(Agency_Id, Site_Id, Service_Id, ServiceOption_PaymentOptions)

test %>%
  splitListToCols(ServiceOption_PaymentOptions)

test %>%
  splitListToCols(ServiceOption_PaymentOptions, cols = FALSE) %>%
  pivot_wider(names_from = ServiceOption_PaymentOptions, values_from = 1, names_prefix = "ServiceOption_PaymentOptions_")
# pivot_wider(names_from = ServiceOption_PaymentOptions, values_from = bool)

test %>%
  splitListToCols(field = ServiceOption_PaymentOptions, cols = FALSE) %>%
  spreadRowsToCols(ServiceOption_PaymentOptions)

test %>%
  splitListToCols(field = ServiceOption_PaymentOptions, cols = TRUE)
```

# Check encodable fields

```{r}
services_original %>% getUniqueVals(ServiceTaxonomy_ModuleServicepost)
services_original %>% getUniqueVals(ServiceOption_PaymentOptions)
services_original %>% getUniqueVals(ServiceOption_PopulationFocus)
services_original %>% getUniqueVals(ServiceOption_LanguagesOtherThanEnglishNew)
services_original %>% getUniqueVals(ServiceOption_Days)
services_original %>% getUniqueVals(ServiceOption_RecordOwner)
services_original %>% getUniqueVals(ServiceOption_MentalHealthConditions)
services_original %>% getUniqueVals(ServiceOption_MedicalConditions)
services_original %>% getUniqueVals(ServiceOption_CounselingTypesOffered)
services_original %>% getUniqueVals(ServiceOption_AgeGroup)
services_original %>% getUniqueVals(ServiceOption_TherapyAndSupportiveApproaches)
services_original %>% getUniqueVals(ServiceOption_IssueOfFocus)
services_original %>% getUniqueVals(ServiceOption_Aoda)
```

# Extract hours of operation data

```{r}
times <- foreach(
  i = 1:nrow(services_original)
  , .combine = rbind
) %do% {
  this <- services_original$ServiceHoursofoperation_ModuleHoursofoperation.open[i] %>% parseTimeString() %>% createTimeStrings()
}
print(times)
```


# Data cleanup

```{r}
services <- services_original %>%
  splitListToCols(ServiceOption_PaymentOptions) %>%
  splitListToCols(ServiceOption_PopulationFocus) %>%
  splitListToCols(ServiceOption_LanguagesOtherThanEnglishNew) %>%
  splitListToCols(ServiceOption_Days) %>%
  splitListToCols(ServiceOption_MentalHealthConditions) %>%
  splitListToCols(ServiceOption_MedicalConditions) %>%
  splitListToCols(ServiceOption_CounselingTypesOffered) %>%
  splitListToCols(ServiceOption_AgeGroup) %>%
  splitListToCols(ServiceOption_TherapyAndSupportiveApproaches) %>%
  splitListToCols(ServiceOption_IssueOfFocus) %>%
  splitListToCols(ServiceOption_Aoda)
```

# Write data

```{r}
write_csv(services, file = "../uwwi_dataset_services_encoded.csv")
```

