---
title: "UWWI Sites cleanup"
output:
  html_notebook:
    code_folding: show
    toc: true
    toc_float: true
---

R code to take the sites dataset and convert columns with lists of values, e.g. ['Sunday', 'Monday'] to long or wide format as desired.

Requires the functions stored separately in `functions.R` to work.

```{r setup}
library(tidyverse)
library(foreach)

source(file = "DataCleaningFunctions.R")

# Hide the dplyr .groups message
options(dplyr.summarise.inform = FALSE)
```

```{r}
sites_original <- read_csv("../uwwi_dataset_sites.csv")
```

```{r}
summary(sites_original)
```

There shouldn't be 0 latitudes and -124 longitudes. Data cleaning issue.

Look at the hours of operation column.

```{r}
hoo <- sites_original %>%
  mutate(hoo = factor(SiteHoursofoperation_ModuleHoursofoperation.open)) %>%
  group_by(hoo) %>%
  summarise(n = n())

print(hoo)
```

Hopefully the 9803 are normal hours, rather than empty hours.

```{r}
hoo %>%
  head(1) %>%
  mutate(hoo = as.character(hoo)) %>%
  unlist()
```

The 9803 don't have hours listed.

```{r}
sites_original %>%
  filter(SiteHoursofoperation_ModuleHoursofoperation.open == "[{'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Sunday', 'start_min': None, 'start_hour': None}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Monday', 'start_min': None, 'start_hour': None}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Tuesday', 'start_min': None, 'start_hour': None}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Wednesday', 'start_min': None, 'start_hour': None}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Thursday', 'start_min': None, 'start_hour': None}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Friday', 'start_min': None, 'start_hour': None}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Saturday', 'start_min': None, 'start_hour': None}]") %>%
  head(50) %>%
  select(AgencySystem_Name, SiteSystem_Active, SiteCustom_NonStandardHoursText, SiteHoursofoperation_ModuleHoursofoperation.open) %>%
  print()
```

Can't realistically clean this. Make a function to parse the hours of operation instead.

```{r}
# Parse field of interest from time string
# Uses regular expressions to pull from the tags and create a valid data frame
parseTimeString <- function(s) {
  # Generate data frame of time parts
  tlist <- list(
     dayOfWeek = str_extract_all(s, "(?<='dayOfWeek': ')[A-z][A-z][A-z]")
    , start_hour = str_extract_all(s, "(?<='start_hour': )(\\[([0-9]+(\\]|,))|(None))")
    , start_min = str_extract_all(s, "(?<='start_min': )(\\[([0-9]+(\\]|,))|(None))")
    , end_hour = str_extract_all(s, "(?<='end_hour': )(\\[([0-9]+(\\]|,))|(None))")
    , end_min = str_extract_all(s, "(?<='end_min': )(\\[([0-9]+(\\]|,))|(None))")
  )
  tdf <-  as.data.frame(tlist)
  names(tdf) <- names(tlist)
  # Convert hour/minute to numeric
  tdf %>%
    mutate(
      across(
          .cols = start_hour:end_min
          , .fns = function(x) {suppressWarnings(stripEnclosingChars(x) %>% as.numeric())}
      )
    ) %>%
    return()
}

# Function to create SunStart, SunEnd, MonStart, MonEnd, etc. strings from output of parseTimeString
createTimeStrings <- function(tdf) {
  # Helper function to format h:mm
  time_pad <- function(h, m) {
    paste0(h, ":", str_pad(m, width = 2, side = "left", pad = "0"))
  }
  # Create new columns
  tdf %>%
    mutate(
      Start = time_pad(start_hour, start_min)
      , End = time_pad(end_hour, end_min)
      , across(
        .cols = Start:End
        , .fns = function(x) {ifelse(x == "NA:NA", NA, x)}
      )
    ) %>%
    select(-ends_with("hour"), -ends_with("min")) %>%
    pivot_wider(names_from = dayOfWeek, values_from = Start:End) %>%
    data.frame() %>%
    return()
}
```

```{r}
# Unit tests
tsample <- "[{'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Sunday', 'start_min': None, 'start_hour': None}, {'day': 1, 'end_min': [0], 'end_hour': [21], 'dayOfWeek': 'Monday', 'start_min': [0], 'start_hour': [8]}, {'day': 2, 'end_min': [0], 'end_hour': [21], 'dayOfWeek': 'Tuesday', 'start_min': [0], 'start_hour': [8]}, {'day': 3, 'end_min': [0], 'end_hour': [21], 'dayOfWeek': 'Wednesday', 'start_min': [0], 'start_hour': [8]}, {'day': 4, 'end_min': [0], 'end_hour': [21], 'dayOfWeek': 'Thursday', 'start_min': [0], 'start_hour': [8]}, {'day': 5, 'end_min': [0], 'end_hour': [21], 'dayOfWeek': 'Friday', 'start_min': [0], 'start_hour': [8]}, {'day': 6, 'end_min': [0], 'end_hour': [21], 'dayOfWeek': 'Saturday', 'start_min': [0], 'start_hour': [8]}]"

tsamp <- parseTimeString(tsample)
print(tsamp)

tsamp %>%
  createTimeStrings()
```

Some have multiple start/stop hours in a single day, annoyingly.

```{r}
parseTimeString("[{'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Sunday', 'start_min': None, 'start_hour': None}, {'day': 1, 'end_min': [0, 0], 'end_hour': [12, 17], 'dayOfWeek': 'Monday', 'start_min': [0, 0], 'start_hour': [8, 13]}, {'day': 2, 'end_min': [0, 0], 'end_hour': [12, 18], 'dayOfWeek': 'Tuesday', 'start_min': [0, 0], 'start_hour': [8, 13]}, {'day': 3, 'end_min': [0, 0], 'end_hour': [12, 18], 'dayOfWeek': 'Wednesday', 'start_min': [0, 0], 'start_hour': [8, 13]}, {'day': 4, 'end_min': [0, 0], 'end_hour': [12, 17], 'dayOfWeek': 'Thursday', 'start_min': [0, 0], 'start_hour': [8, 13]}, {'day': 5, 'end_min': [0], 'end_hour': [12], 'dayOfWeek': 'Friday', 'start_min': [0], 'start_hour': [8]}, {'day': -1, 'end_min': None, 'end_hour': None, 'dayOfWeek': 'Saturday', 'start_min': None, 'start_hour': None}]
")
```

```{r}
times <- foreach(
  i = 1:nrow(sites_original)
  , .combine = rbind
) %do% {
  this <- sites_original$SiteHoursofoperation_ModuleHoursofoperation.open[i] %>% parseTimeString() %>% createTimeStrings()
}
print(times)
```

```{r}
sites <- sites_original %>%
  cbind(times)
```

Save data.

```{r}
write_csv(sites, file = "../uwwi_dataset_sites_encoded.csv")
```

